<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Cleaning data using GBIF issues}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
	comment = "#>",
	collapse = TRUE,
	warning = FALSE,
	message = FALSE
)
```

Cleaning data using GBIF issues
======

`rgbif` now has the ability to clean data retrieved from GBIF based on GBIF issues. These issues are returned in data retrieved from GBIF, e.g., through the `occ_search()` function. Inspired by `magrittr`, we've setup a workflow for cleaning data based on using the operator `%>%`. You don't have to use it, but as we show below, it can make the process quite easy.

Note that you can also query based on issues, e.g., `occ_search(taxonKey=1, issue='DEPTH_UNLIKELY')`. However, we imagine it's more likely that you want to search for occurrences based on a taxonomic name, or geographic area, not based on issues, so it makes sense to pull data down, then clean as needed using the below workflow with `occ_issues()`.

Note that `occ_issues()` only affects the data element in the gbif class that is returned from a call to `occ_search()`. Maybe in a future version we will remove the associated records from the hierarchy and media elements as they are remove from the data element.

`occ_issues()` also works with data from `occ_download()`.

## Get rgbif

Install from CRAN

```{r install, eval=FALSE}
install.packages("rgbif")
```

Or install the development version from GitHub

```{r installgh, eval=FALSE}
devtools::install_github("ropensci/rgbif")
```

Load rgbif

```{r}
library('rgbif')
```

## Get some data

Get taxon key for _Helianthus annuus_

```{r}
(key <- name_suggest(q='Helianthus annuus', rank='species')$key[1])
```

Then pass to `occ_search()`

```{r}
(res <- occ_search(taxonKey=key, limit=100))
```

## Examine issues

The dataset `gbifissues` can be retrieved using the function `gbif_issues()`. The dataset's first column `code` is a code that is used by default in the results from `occ_search()`, while the second column `issue` is the full issue name given by GBIF. The third column is a full description of the issue.

```{r}
head(gbif_issues())
```

You can query to get certain issues

```{r}
gbif_issues()[ gbif_issues()$code %in% c('cdround','cudc','gass84','txmathi'), ]
```

The code `cdround` represents the GBIF issue `COORDINATE_ROUNDED`, which means that

> Original coordinate modified by rounding to 5 decimals.

The content for this information comes from [http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html](http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html).

## Parse data based on issues

Now that we know a bit about GBIF issues, you can parse your data based on issues. Using the data generated above, and using the function `%>%` imported from `magrittr`, we can get only data with the issue `gass84`, or `GEODETIC_DATUM_ASSUMED_WGS84` (Note how the records returned goes down to 98 instead of the initial 100).

```{r}
res %>%
  occ_issues(gass84)
```

Note also that we've set up `occ_issues()` so that you can pass in issue names without having to quote them, thereby speeding up data cleaning.

Next, we can remove data with certain issues just as easily by using a `-` sign in front of the variable, like this, removing data with issues `depunl` and `mdatunl`.

```{r}
res %>%
  occ_issues(-depunl, -mdatunl)
```

## Expand issue codes to full names

Another thing we can do with `occ_issues()` is go from issue codes to full issue names in case you want those in your dataset (here, showing only a few columns to see the data better for this demo):

```{r}
out <- res %>% occ_issues(mutate = "expand")
head(out$data[,c(1,5)])
```


## Add columns

Sometimes you may want to have each type of issue as a separate column.

Split out each issue type into a separate column, with number of columns equal to number of issue types

```{r}
out <- res %>% occ_issues(mutate = "split")
head(out$data[,c(1,5:10)])
```

## Expand and add columns

Or you can expand each issue type into its full name, and split each issue into a separate column.

```{r}
out <- res %>% occ_issues(mutate = "split_expand")
head(out$data[,c(1,5:10)])
```

## Wrap up

We hope this helps users get just the data they want, and nothing more. Let us know if you have feedback on data cleaning functionality in `rgbif` at _info@ropensci.org_ or at [https://github.com/ropensci/rgbif/issues](https://github.com/ropensci/rgbif/issues).
