<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Tutorial for using the old GBIF API}
-->

rgbif tutorial
--------------

*Note: this vignette works with the current version on CRAN that works with the old GBIF API.*

The `rgbif` package interacts with the API services of the Global Biodiversity Information Facility [GBIF](http://www.gbif.org/). GBIF currently holds 377,177,914 indexed records, ~10K datasets, and 419
publishers (i.e., dataset submitters).

This tutorial will go through three use cases to demonstrate the kinds of things possible in `rgbif`.

* Counts taxon concept records matching a range of filters.
* Returns summary counts of occurrence records by one-degree cell.
* occurrencelist searches for taxon concept records matching a range
of filters.
* densitylist gets density of occurrence records by one-degree cell.
* Search by taxon to retrieve number of records in GBIF.

## Install and load package from GitHub

```{r install, eval=FALSE}
install.packages("rgbif")
```

```{r load, message=FALSE, warning=FALSE}
library(rgbif)
```

## Counts taxon concept records matching a range of filters.

```{r occurrencecount1, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
occurrencecount(scientificname = "Helianthus annuus", coordinatestatus = TRUE, year = 2005, maxlatitude = 20)
```

Count many taxa

```{r occurrencecount2, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
lapply(c("Helianthus debilis", "Abies procera", "Astragalus"), function(x) occurrencecount(scientificname = x, coordinatestatus = TRUE))
```

## Return summary counts of occurrence records by one-degree cell for a single taxon, country, dataset, data publisher or data network

```{r densitylist, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
out <- densitylist(originisocountrycode = "CA")
head(gbifdata(out))
```

## Occurrencelist searches for taxon concept records matching a range of filters.

A simple example

```{r occurrencelist, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
dat <- occurrencelist(scientificname = "Accipiter erythronemius", coordinatestatus = TRUE, maxresults = 10)
gbifdata(dat)
```

Search for many species and make a map

```{r occurrencelist_many, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
splist <- c("Accipiter erythronemius", "Junco hyemalis", "Aix sponsa")
out <- occurrencelist_many(splist, coordinatestatus = TRUE, maxresults = 20)
gbifmap_list(out)
```

## densitylist provides access to records showing the density of occurrence records from the GBIF Network by one-degree cell.

A simple example

```{r densitylist2, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
out <- densitylist(originisocountrycode="US")
gbifmap_dens(out)
```

## Search by taxon to retrieve number of records in GBIF.

```{r taxoncount1, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
taxoncount("Puma concolor")
```

```{r taxoncount2, message=FALSE, warning=FALSE, comment=NA, cache=FALSE}
taxoncount("Helianthus annuus")
```