<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->

rgbif vignette - Seach and retrieve data from the Global Biodiverity Information Facilty (GBIF)
======

### About the package

`rgbif` is an R package to search and retrieve data from the Global Biodiverity Information Facilty (GBIF). `rgbif` wraps R code around the [GBIF API][gbifapi] to allow you to talk to the BISON database from R. 

********************

#### Install rgbif

```{r install, comment=NA, warning=FALSE}
# install.packages("devtools"); library(devtools); install_github("rbison", "ropensci")
library(rgbif); library(XML); library(RCurl); library(plyr); library(ggplot2); library(maps)
```

********************

#### Get a list of the data networks in GBIF - and you can use the networkkey number to seach for occurrences for the specific provider in other functions

```{r networks, comment=NA, warning=FALSE}
# Test the function for a few networks
networks(maxresults=5)

# By name
networks('ORNIS')
```

********************

#### Get a list of the data providers in GBIF - and you can use the dataproviderkey number to seach for occurrences for the specific provider in other functions

```{r providers, comment=NA, warning=FALSE}
# Test the function for a few providers
providers(maxresults=5)

# By data provider name
providers('University of Texas-Austin')
```

********************

#### Get a list of the data resources in GBIF - and you can use the resourcekey number to seach for occurrences for the specific resource in other functions

```{r resources, comment=NA, warning=FALSE}
# Test the function for a few resources
resources(maxresults=5)

# By name
head(resources('Flora'))
```

********************

#### Get number of occurrences for a set of search parameters
```{r occurrencecount, comment=NA, warning=FALSE}
occurrencecount(scientificname = 'Accipiter erythronemius', coordinatestatus = TRUE)
occurrencecount(scientificname = 'Helianthus annuus', coordinatestatus = TRUE, year=2005, maxlatitude=20)
```

********************

#### Get possible values to be used in taxonomic rank arguments in functions

```{r taxrank, comment=NA, warning=FALSE}
taxrank()
```

********************

#### Seach by taxon to retrieve number of records per taxon found in GBIF
```{r taxoncount, comment=NA, warning=FALSE}
taxoncount(scientificname = 'Puma concolor')
taxoncount(scientificname = 'Helianthus annuus')
taxoncount(rank = 'family')
```

********************

#### Get taxonomic information on a specific taxon or taxa in GBIF by their taxon concept keys

```{r taxonget, comment=NA, warning=FALSE}
(out <- taxonsearch(scientificname = 'Puma concolor'))
```

********************

#### Search for taxa in GBIF

```{r taxonsearch, comment=NA, warning=FALSE}
taxonsearch(scientificname = 'Puma concolor', rank="species", maxresults=10)
taxonsearch(scientificname = 'Puma concolor', rank="species", dataproviderkey=1)
```

********************

#### Get data for a single occurrence. Note that data is returned as a list, so you have to convert to a data.frame, etc. as you wish
```{r occurrenceget, comment=NA, warning=FALSE}
occurrenceget(key = 13749100)$dataProvider$dataResources$dataResource$occurrenceRecords$TaxonOccurrence[1:10]
```

********************

```{r occurrencelist, comment=NA, warning=FALSE}
out <- occurrencelist(scientificname = 'Puma concolor', coordinatestatus=TRUE, maxresults=20)
```

Note that the default object printed from a call to `occurrencelist` is a list that contains:

+ NumberFound: number of occurrences found in search results.
+ TaxonNames: Unique list of taxonomic names in search results.
+ Coordinates: Min and max latitude and longitude of all occurrences.
+ Countries: Countries contained in results set.

```{r defaultoccurencelist, comment=NA, warning=FALSE}
out
```

Where do you get data after a call to the `occurrencelist` function? This is where `gbifdata` comes in. By default a call to `gbifdata` prints a minimal data.frame with just rows *name*, *latitude*, and *longitude*.

```{r minimaltrue, comment=NA, warning=FALSE}
gbifdata(out)
```

Though you can get more detailed data by calling *minimal=FALSE*.

```{r minimalfalse, comment=NA, warning=FALSE}
head( gbifdata(out, minimal=FALSE)[,1:6] )
```

And you can get all possible data by specifying *format=darwin*.

```{r occurrencelistdarwin, comment=NA, warning=FALSE}
out <- occurrencelist(scientificname = 'Puma concolor', coordinatestatus=TRUE, format="darwin", maxresults=20)
head( gbifdata(out, minimal=FALSE)[,1:6] )
```

********************

#### Maps

```{r gbifmap1, comment=NA, warning=FALSE}
splist <- c('Accipiter erythronemius', 'Junco hyemalis', 'Aix sponsa', 'Ceyx fallax', 'Picoides lignarius', 'Campephilus leucopogon')
out <- occurrencelist_many(splist, coordinatestatus = TRUE, maxresults = 20)
gbifmap_list(out)
```

Another example, setting scientificname="*" so we just get any species, and then mapping points only within the state of Texas in the US.

```{r gbifmap2, comment=NA, warning=FALSE}
out <- occurrencelist(scientificname="*", minlatitude=30, maxlatitude=35, minlongitude=-100, maxlongitude=-95, coordinatestatus = TRUE, maxresults = 200)
gbifmap_list(input=out, mapdatabase="state", region="texas", geom=geom_jitter, jitter=position_jitter(width = 0.3, height = 0.3))
```

********************

[gbifapi]: http://data.gbif.org/tutorial/services